---
title: "Agricultural Trends"
author: "Hadrien@rstudio.com"
date: "3/27/2019"
output:
  html_document: default
params:
  country:
    choices:
    - United States of America
    - Mexico
    - Canada
    - Saudi Arabia
    - Brazil
    - France
    - Italy
    input: select
    label: Country
    value: United States of America
  crop:
    choices: [Wheat, Rice, Maiz, Barley, Apples, "Beans, dry", Sugar cane, Carrots and turnips, Bananas]
    input: select
    label: Crop
    value: Wheat
---

# Overview 

This analysis aims to better understand agricultural production trends in order to better guide decisions related to the manufacture of farming equipment to ensure future food security in different geographies.


**Data:** Data for this analysis was obtained from the [Food and Agricultural Organization of the United Nations](http://www.fao.org/faostat/en/#data/QC) website on April 1st, 2019. Data was downloaded as the bulk download "All Data".

**Using this report** 

To view this report, go to the [report URL](https://colorado.rstudio.com/rsc/agrilculture-trends-paramaterized-with-inputs/) on RStudio Connect. On the left hand side of the report, select your desired input paramaters and click run. A report can be generated for any country & crop combination. This report was designed for the following users: 

- Marketing managers: explore production trends to plan which markets to invest in and create campaigns for. 
- Product managers: identify which product lines to invest in to help farmers overcome challenges such as droughts and desertification and meet the demand of an ever increasing population.

**Contributing** 

To contribute to this project, fork this report's repository on [github](https://github.com/hdykiel/paramaterized-report-agriculture/commits/master) and create a new branch using the format "your-name-feature-name" and submit a pull request.


```{r setup, include=FALSE}
# set default settings for all code chunks
knitr::opts_chunk$set(message=FALSE, warning=FALSE, echo=FALSE)

# load required packages
library(tidyverse)
library(readr)
library(choroplethr)
library(choroplethrMaps)
library(blastula)
```


```{r tidy-data, include=FALSE}

######################
# load and tidy data #
######################


# load data
df <- read_csv("../data/Production_Crops_E_All_Data_NOFLAG.csv")

# transform from wide to long format
df <- df %>% gather(Year, Value, Y1961:Y2017)

# clean up year
df$Year <-  df$Year %>% str_remove("Y") %>% as.integer()

# remove aggregates and yield
df <- df %>% 
  filter(Element == "Production", !grepl("Total|Primary|Rice Milled Eqv", Item))


```



```{r top-crops, eval=FALSE, echo=FALSE, message=FALSE}

######################################
# Top production by crop (worldwide) #
######################################

# summarize by food item
df_total_production <-  df %>%
  group_by(Item) %>%
  summarise(tonnes_produced = sum(na.omit(Value)))

# total production by crop barplot
ggplot <- df_total_production %>%
  top_n(20, tonnes_produced) %>%
  ggplot(aes(reorder(Item,-tonnes_produced), tonnes_produced)) + geom_bar(stat =
                                                                            "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```


```{r country-stats}
data("country.map") # load data set
country.data <- country.map %>%
  select(region, type, pop_est, region_wb, economy) %>% # keep only certain metrics
  distinct() # retain unique rows only

# keep stats for selected countries only
country <- params$country
crop <- params$crop

stats <- country.data %>% filter(region==str_to_lower(params$country))
country_url <- paste0("https://en.wikipedia.org/wiki/", str_to_title(country))
```




```{r plot1-choropleth, echo=FALSE}

# summarize total production by country
df_total_by_country <- df %>%
  group_by(Area) %>%
  summarize(value = sum(na.omit(Value))) %>%
  mutate(region = str_to_lower(Area)) %>% 
  select(region, value) # reorganize columns, drop Area

# filter any country name not found in choroplethrMaps::country.map data frame
df_total_by_country <- df_total_by_country %>%
  filter(region %in% country.map$region)

# production by country choropleth
plot1 <- country_choropleth(df_total_by_country)


```


 
```{r plot2-production,echo=FALSE}

# plot production by crop and country
plot2 <- df %>%
  filter(Item == crop, Area == country) %>%
  ggplot(aes(Year, Value)) + geom_point() +
  geom_smooth() +
  ggtitle(label = paste("Annual Production (in tonnes) of", params$crop, "in", params$country))
```


### Worldwide production of `r params$crop`


```{r plot1, fig.width=9}
plot1
```

 

<div style="float:right; margin-left: 10%;">

```{r plot2, fig.width=6}
plot2
```

</div> 

### Production trend (`r params$country`)

 
[`r str_to_title(stats$region)`](`r country_url`) is a country in `r stats$region_wb` with an estimated population of `r stats$pop_est`. Its government system is `r stats$type` and is categorized as having a "`r stats$economy`" economy.



```{r email-options}

# Set e-mail subject
rmarkdown::output_metadata$set(rsc_email_subject = "My awesome e-mail subject")

# Construct e-mail subject
subject <- rmarkdown::output_metadata$get("rsc_email_subject") # retrieve current email subject
difference <- (1200-1000)/1000 # some metric
rmarkdown::output_metadata$set(rsc_email_subject = paste0(subject," - ", difference*100, "%", " change!"))

# Suppress e-mail
if (difference < .1) {
  rmarkdown::output_metadata$set(rsc_suppress_email_scheduled = TRUE)
}
```




```{r compose-email}

# Create a simple email message using
# Markdown formatting with the blastula package
library(blastula)

email <-
  compose_email(
  body = "
  Time generated: {Sys.time()}

  ## {params$country} {params$crop} production

  We can use Markdown formatting \\
  to **embolden** text or to add \\
  *emphasis*.
  
  {add_ggplot(plot_object = plot2, width = 8, height = 5)}

  Please contact the reporting team with any questions about this report.")

# The email message can always be
# previewed using `preview_email()`
# preview_email(email = email)

# Give the HTML email data to RStudio Connect.
rmarkdown::output_metadata$set(rsc_email_body_html = email$html_str)
rmarkdown::output_metadata$set(rsc_email_images = email$images)


```

