---
title: "State Profitability EDA"
author: "Hadrien@rstudio.com"
date: "9/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
```

In this script we performed some preliminary exploratory data analysis to better understand the data. Unlike `01_simulate_data.Rmd`, this script does not need to be executed in order to generate the final state profitibility report. It may, however, be useful in answering questions which arise from the report.

We first began by generating a few different choropleth maps of the US showing loss ratio. We used the code provided on the webpage ttps://socviz.co/maps.html as a starting point.

```{r data-prep}
# load policy data
dat <- read_csv("data.csv")

# load map data
library(maps)
us_states <- map_data("state")
head(us_states)


```

```{r visualize, include = TRUE}

# a basic map
p1 <- ggplot(data = us_states,
            mapping = aes(x = long, y = lat,
                          group = group, fill = region))

p1 + geom_polygon(fill = "white", color = "black")

# fill by LR
us_states_lr <- left_join(us_states, dat, by = c("region"= "state")) # join data sets

p2 <- ggplot(data = us_states_lr,
            aes(x = long, y = lat,
                group = group, fill = loss_ratio))


p2 + geom_polygon(color = "gray90", size = 0.1) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) 

# add a fill gradient
(p3 <- ggplot(data = us_states_lr, aes(x = long, y = lat, group = group, fill = loss_ratio)) +
  geom_polygon(color = "gray90", size = 0.1) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  scale_fill_gradient2(midpoint = 0.5, high = scales::muted("red"), low = scales::muted("blue") )
)




```

