---
title: "Agricultural Trends"
author: "Hadrien@rstudio.com"
date: "3/27/2019"
output: html_document
---

# Overview 

This analysis aims to better understand agricultural production trends in order to better guide decisions related to the manufacture of farming equipment to ensure future food security in different geographies.


**Data:** Data for this analysis was obtained from the [Food and Agricultural Organization of the United Nations](http://www.fao.org/faostat/en/#data/QC) website on April 1st, 2019. Data was downloaded as the bulk download "All Data".

**Using this report** 

To view this report, go to the report URL on RStudio Connect. On the left hand side of the report, select your desired input paramaters and click run. A report can be generated for any country & crop combination. This report was designed for the following users: 

- Marketing managers: explore production trends to plan which markets to invest in and create campaigns for. 
- Product managers: identify which product lines to invest in to help farmers overcome challenges such as droughts and desertification and meet the demand of an ever increasing population.

**Contributing** 

To contribute to this project, fork this report's repository on [github](https://github.com/hdykiel/paramaterized-report-agriculture/commits/master) and create a new branch using the format "your-name-feature-name" and submit a pull request.


```{r setup, include=FALSE}
# set default settings for all code chunks
knitr::opts_chunk$set(message=FALSE, warning=FALSE, echo = FALSE)

# load required packages
library(tidyverse)
library(readr)
library(choroplethr)
library(choroplethrMaps)

```


```{r tidy-data, include=FALSE}
######################
# load and tidy data #
######################

# load data
df <- read_csv("../data/Production_Crops_E_All_Data_NOFLAG.csv")

# transform from wide to long format
df <- df %>% gather(Year, Value, Y1961:Y2017)

# clean up year
df$Year <-  df$Year %>% str_remove("Y") %>% as.integer()

# remove aggregates and yield
df <- df %>% 
  filter(Element == "Production", !grepl("Total|Primary|Rice Milled Eqv", Item))


```



```{r top-crops, eval=FALSE, echo=FALSE, message=FALSE}
######################################
# Top production by crop (worldwide) #
######################################

# summarize by food item
df_total_production <-  df %>%
  group_by(Item) %>%
  summarise(tonnes_produced = sum(na.omit(Value)))

# total production by crop barplot
ggplot <- df_total_production %>%
  top_n(20, tonnes_produced) %>%
  ggplot(aes(reorder(Item,-tonnes_produced), tonnes_produced)) + geom_bar(stat =
                                                                            "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```


```{r country-stats}
data("country.map") # load data set
country.data <- country.map %>%
  select(region, type, pop_est, region_wb, economy) %>% # keep only certain metrics
  distinct() # retain unique rows only


```




```{r plot1-choropleth, echo=FALSE}

# summarize total production by country
df_total_by_country <- df %>%
  group_by(Area) %>%
  summarize(value = sum(na.omit(Value))) %>%
  mutate(region = str_to_lower(Area)) %>% 
  select(region, value) # reorganize columns, drop Area

# filter any country name not found in choroplethrMaps::country.map data frame
df_total_by_country <- df_total_by_country %>%
  filter(region %in% country.map$region)

# production by country choropleth
plot1 <- country_choropleth(df_total_by_country)


```


 
```{r plot2-production,echo=FALSE}

# plot production by crop and country
plot2 <- df %>%
  filter(Item == "Wheat", Area == "United States of America") %>%
  ggplot(aes(Year, Value)) + geom_point() +
  geom_smooth() +
  ggtitle(label = "Annual Production (in tonnes) of Wheat in USA")
```


### Worldwide production of Wheat


```{r plot1, fig.width=9}
plot1
```

 

<div style="float:right; margin-left: 10%;">

```{r plot2, fig.width=6}
plot2
```

</div> 

### Production trend USA

Insert markdown text here...


